---
title: "Group Presentation"
author: 
  - "Dea Avega Editya"
  - "Siyi Li"
  - "Xitong He"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    seal : false
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse
```{r setup, include=FALSE,warning=FALSE,message=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(ggplot2)
library(plotly)
library(lubridate)
library(fontawesome)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#3e8cb4",
  inverse_header_color = "#FFFFFF"
)
```


# Group Presentation

### Team: Pandorea
--

#### Team Members:

- `r fa("grin", fill = "yellow")` "Dea Avega Editya"

- `r fa("grin-stars", fill = "yellow")` "Siyi Li"

- `r fa("smile-wink", fill = "yellow")` "Xitong He"

--

.center[
# "Make a Plot Bad Again!" `r fa("fist-raised", fill = "white")`
]
---

# Data collection `r fa("database")`

Data source: [dugong](https://github.com/etc5521-2020/assignment-1-dugong) repo

#### Reading the dataset

```{r load-data,echo=TRUE, eval = TRUE, warning=FALSE,message=FALSE}
mortgage_rates <- readr::read_csv(here::here("data/mortgage_rates.csv"))
state_hpi <-readr::read_csv(here::here("data/state_hpi.csv"))

```

```{r , echo = FALSE, eval = TRUE, warning=FALSE,message=FALSE}
state_hpi <- state_hpi %>%
  mutate(date = ymd(year*10000+month*100+1))

nation_hpi <- state_hpi %>%
  dplyr::select("date","year","month","us_avg") %>%
  distinct()
recession_dates <- readr::read_csv(here::here("data/recession_dates.csv")) 
recession_clean <- recession_dates %>% 
  mutate(duration_months = substring(duration_months, 3),
         period_range = substring(period_range, 5),
         time_since_previous_recession_months = substring(time_since_previous_recession_months, 4),
         peak_unemploy_ment = substring(peak_unemploy_ment, 5),
         period_range = case_when(name == "Great Depression" ~ "Aug 1929–Mar 1933",
                                  name == "Great Recession" ~ "Dec 2007–June 2009",
                                  name == "COVID-19 recession" ~ "Feb 2020–present",
                                  TRUE ~ period_range)) %>% 
mutate(peak_unemploy_ment = recode(peak_unemploy_ment,
                                   "5.2%[57](1946)" = "5.2%(1946)",
                                   "17.8%[50]–\n19.0%(1938)[57]" = "17.8%–19.0%(1938)",
                                   "21.3%(1932)[50]–  24.9%(1933)[51]"= "21.3%(1932)–24.9%(1933)",
                                   "10.0%(October 2009)[76]" = "10.0%(October 2009)")) 
```

```{r , echo =FALSE, eval = TRUE, warning=FALSE,message=FALSE}
recession_dat <- recession_clean  %>% 
  separate(period_range, into = c("start", "end"), sep = "–") %>% 
  mutate(start = as.Date(parse_date_time(start, "%b %Y")), 
         end = as.Date(parse_date_time(end, "%b %Y"))) 
```

---
# A Brief Story about the Original Plot 

##### Mortgage Rate vs Housing Price Index Plot 

.pull-left[
```{r ori,echo=FALSE,message=FALSE,warning=FALSE,fig.width=9,fig.height=9}
hpi_mort <- ggplot()+ 
  geom_line(data=mortgage_rates, 
            aes(x=date, y=fixed_rate_30_yr*10),
            color = "#A73737",
            size = 0.8) +
  geom_line(data=nation_hpi, 
            aes(x=date, y=us_avg),
            color="#9BCD9B", 
            size = 0.8)+
  annotate("text", x=as.Date("2012-05-01"), y=110, label="HPI", color = "#9BCD9B", size = 4) +
  annotate("text", x=as.Date("1987-11-27"), y=175, label="Mortgage Rates", color = "#A73737", size = 4) +
  scale_y_continuous(sec.axis = sec_axis(~ . *0.1, name = "Mortgage Rates"))+
  theme_light()+
  xlim(ymd(19750101), NA) +
  labs(x="Year", y="US Average HPI")
  
mortgage_recession <- hpi_mort +
  geom_rect(data = recession_dat, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
            alpha = 0.25, 
            fill = "#7AC5CD") +
  ggtitle("Recession, HPI and Mortgage Rates")+
  theme_light()

mortgage_recession
```
]

--
.pull-right[
* Created by Dugong team
* Showing movement of mortgage rate (fixed rate 30-year tenor) versus Housing Price Index in United States of America
* Data covers period from 1971 to 2018 in monthly basis
* Its purpose is to explore relationship between these two variables  
]

---

# Why is it Great `r fa("thumbs-up", fill = "brown")`
.pull-left[
```{r plotly,echo=FALSE,message=FALSE,warning=FALSE,fig.width=9,fig.height=9}
hpi_mort <- ggplot()+ 
  geom_line(data=mortgage_rates, 
            aes(x=date, y=fixed_rate_30_yr*10),
            color = "#A73737",
            size = 0.8) +
  geom_line(data=nation_hpi, 
            aes(x=date, y=us_avg),
            color="#9BCD9B", 
            size = 0.8)+
  annotate("text", x=as.Date("2012-05-01"), y=110, label="HPI", color = "#9BCD9B", size = 4) +
  annotate("text", x=as.Date("1987-11-27"), y=175, label="Mortgage Rates", color = "#A73737", size = 4) +
  scale_y_continuous(sec.axis = sec_axis(~ . *0.1, name = "Mortgage Rates"))+
  theme_light()+
  xlim(ymd(19750101), NA) +
  labs(x="Year", y="US Average HPI")
  
mortgage_recession <- hpi_mort +
  geom_rect(data = recession_dat, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
            alpha = 0.25, 
            fill = "#7AC5CD") +
  ggtitle("Recession, HPI and Mortgage Rates")+
  theme_light()

mortgage_recession
```
]

.pull-right[
- Using line plot to construct the time series via the `geom_line` function 
- Scaling up mortgage rate (fixed_rate_30_yr) by multiplying it with 10 

- having the same starting date 1975/01/01 (by setting xlim for both)

- different color to distinguish the line trend

- adding the annotation could be easy to understand which one stands for HPI or Mortgage Rates

]
---
.center[

# Let's Make It Ugly!

]
--
.center[
# From Hero to Zero
]
--
.pull-left[
<img src = "https://upload.wikimedia.org/wikipedia/en/9/9a/Trollface_non-free.png" width = 50%>

]

.pull-right[
<img src = "https://assets.stickpng.com/images/580b585b2edbce24c47b2a2f.png" width = 50%>

]
---
# Bad Plot `r fa("frown", fill = "red")`

.pull-left[
```{r bad, message=FALSE,echo=FALSE,warning=FALSE}
ggplot()+
  geom_point(data=mortgage_rates,
           aes(x=date, y=fixed_rate_30_yr, size= fixed_rate_30_yr),alpha=0.2) +
  geom_point(data=nation_hpi,
           aes(x=date, y=us_avg, size=us_avg),alpha=0.2)+
  theme_bw()+
  ylab("rate")+
    geom_rect(data = recession_dat, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),fill="grey") +
  ggtitle("Recession, HPI and Mortgage Rates")
```
]

.pull-right[

- Use `fixed_rate_30_yr` for y value instead of `fixed_rate_30_yr*10` to draw the time series line.

- Remove the `xlim` function.

- Remove the scales for continuous y aesthetics in secondary axis.

- Use `geom_point` function instead of `geom_line`

]
---
# Bad Plot (Cont) `r fa("frown", fill = "red")`

.pull-left[
```{r bad-cont, message=FALSE,echo=FALSE,warning=FALSE}
ggplot()+
  geom_point(data=mortgage_rates,
           aes(x=date, y=fixed_rate_30_yr, size= fixed_rate_30_yr),alpha=0.2) +
  geom_point(data=nation_hpi,
           aes(x=date, y=us_avg, size=us_avg),alpha=0.2)+
  theme_bw()+
  ylab("rate")+
    geom_rect(data = recession_dat, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),fill="grey") +
  ggtitle("Recession, HPI and Mortgage Rates")
```
]

.pull-right[

- Remove the color palette in two line trends and recession period

- Remove the annotation details in mortgage rate and hpi line trends

- Add relative size for both variables

- Reduce transparency and date rectangles
]


---
## Disadvantage in bad plot `r fa("times-circle", fill = "red")`

- Without the scale(*10), graphs are hard to compare, since the size of the values are different.

- Without the color and annotation, it is hard to distinguish the two rates. 

- Without 'xlim', the datasets start at different time. 

- Without secondary axis, it is difficult to get the rate of another one because of quite different scales.

- Reduce the value of 'alpha' in 'geom_rect', the graphs of rates are not continue anymore since the rectangles cover it.

- Reduce the size of point width, the graph doesn't seem that obvious

- Lots of unrelated duration of the recession in the graph bofore 1970.


---

# Conclusion `r fa("hand-holding-heart", fill = "green")`

* Line plot is the best option for time series
* Make variables comparable by using equal scale
* Add relative size parameter to the plot has not really obvious (and maybe be useless) when observations are large,  
* Differentiate colour of 2 variables can ease reader to understand the plot.