---
title: "Group Presentation"
author: 
  - "Dea Avega Editya"
  - "Siyi Li"
  - "Xitong He"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    seal : false
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse
```{r setup, include=FALSE,warning=FALSE,message=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(ggplot2)
library(plotly)
library(lubridate)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#3e8cb4",
  inverse_header_color = "#FFFFFF"
)
```


# Group Presentation

### Team: Pandorea
--

#### Team Members:

- `r icon::fontawesome("grin")` "Dea Avega Editya"

- `r icon::fontawesome("grin-stars")` "Siyi Li"

- `r icon::fontawesome("smile-wink")` "Xitong He"
---

# Data collection `r icon::fontawesome("database")`

Loading the data from the [dugong](https://github.com/etc5521-2020/assignment-1-dugong) repo:

```{r load-data,warning=FALSE,message=FALSE}
mortgage_rates <- readr::read_csv(here::here("data/mortgage_rates.csv"))
state_hpi <-readr::read_csv(here::here("data/state_hpi.csv"))

state_hpi <- state_hpi %>%
  mutate(date = ymd(year*10000+month*100+1))

nation_hpi <- state_hpi %>%
  dplyr::select("date","year","month","us_avg") %>%
  distinct()
```

---

# Original plot `r icon::fontawesome("thumbs-up")`
.pull-left[
```{r plotly,echo=FALSE,message=FALSE,fig.width=6}
hpi_mort <- ggplot()+
  geom_line(data=mortgage_rates,
            aes(x=date, y=fixed_rate_30_yr*10),
            color = "#A73737",
            size = 0.8) +
  geom_line(data=nation_hpi,
            aes(x=date, y=us_avg),
            color="#9BCD9B",
            size = 0.8)+
  annotate("text", x=as.Date("2012-05-01"), y=110, label="HPI", color = "#9BCD9B", size = 4) +
  annotate("text", x=as.Date("1987-11-27"), y=175, label="Mortgage Rates", color = "#A73737", size = 4) +
  scale_y_continuous(sec.axis = sec_axis(~ . *0.1, name = "Mortgage Rates"))+
  theme_light()+
  xlim(ymd(19750101), NA) +
  labs(x="Year", y="US Average HPI") +
  ggtitle("The Changes between Mortgage rates and HPI")

ggplotly(hpi_mort)
```
]

   
.pull-right[
- using the `fixed_rate_30_yr` in `mortgage_rates` dataset and `us_avg` in `nation_hpi` dataset to construct the time series line trend via the `geom_line` function 

- different color to distinguish the line trend

- adding the annotation could be easy to understand which one stands for HPI or Mortgage Rates

]
---


# Bad Plot `r icon::fontawesome("frown")`

.pull-left[
```{r bad, message=FALSE,echo=FALSE}
ggplot()+
  geom_smooth(data=mortgage_rates,
           aes(x=date, y=fixed_rate_30_yr),size=0.1) +
  geom_smooth(data=nation_hpi,
           aes(x=date, y=us_avg),size=0.1)+
  theme_dark()+
  ggtitle("The Changes between Mortgage rates and HPI")+
  ylab("rate")
```
]

.pull-right[

- Remove the `ggploty` function.

- Use `fixed_rate_30_yr` for y value instead of `fixed_rate_30_yr*10` to draw the time series line.

- Remove the `xlim` function.

- Use `geom_smooth` function instead of `geom_line`

- Remove the y axis continuous scale

- Remove the color palette in two line trends

]
---
# Bad Plot (Cont) `r icon::fontawesome("frown")`

.pull-left[
```{r bad-cont, message=FALSE,echo=FALSE}
ggplot()+
  geom_smooth(data=mortgage_rates,
           aes(x=date, y=fixed_rate_30_yr),size=0.1) +
  geom_smooth(data=nation_hpi,
           aes(x=date, y=us_avg),size=0.1)+
  theme_dark()+
  ggtitle("The Changes between Mortgage rates and HPI")+
  ylab("rate")
```
]

.pull-right[
- Remove the annotation details in mortgage rate and hpi line trends

- Remove the scales for continuous y aesthetics in secondary axis.

- Changed the them background from `theme_light()` to `theme_dark`
]
---
## Disadvantage in bad plot `r icon::fontawesome("times-circle")`

- The size of two values are different, hard to compare how the changes between mortgage rates and hpi.

- The starting date is different in two dataset, removing the `xlim` could make the date variable inconsistency.

- Removing the color and annotation could mislead the mortgage rate and hpi. 

- Using the `geom_smooth` is not appropriate in this case as it aids for in the presence of overplotting.

- Not interactive any more which is not clear for each time line.

---

```{r}
ggplot()+
  geom_smooth(data=mortgage_rates,
           aes(x=date, y=fixed_rate_30_yr)) +
  geom_smooth(data=nation_hpi,
           aes(x=date, y=us_avg))+
  theme_bw()+
  ggtitle("The Changes between Mortgage rates and HPI")+
  ylab("rate")
```


## Disadvantage in bad plot `r icon::fontawesome("times-circle")`

- without the scale(*10), graphs are hard to compare, since the size of the values are different.

- without the color and text, it is not readable. 

- without the size, the width of line is bigger and hard to see the CI. 

- without 'xlim', the datasets start at different time. 

- with 'geom_smooth', miss lots of details
